---
title: "datavis"
author: "Zonghong Yu"
date: "2023-09-18"
output: html_document
---

```{r}
library(reticulate)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(lubridate)
library(plotly)
library(TTR) # For the SMA function

```


```{r, echo=FALSE, results='hide',warning=FALSE,message=FALSE}
df <- read.csv("../../Dataset/project/household_saving.csv")
df$DATE <- as.Date(df$DATE)
head(df)
```

```{r,warning=FALSE,echo=FALSE}
library(ggplot2)


ggplot(data = df, aes(x = DATE, y = W398RC1A027NBEA)) +
  geom_line(color = "DarkViolet") +
  labs(title = "Time Series Plot of Household Income",
       x = "Date", 
       y = "The Household Income") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))


```






### ACF & PACF
```{r, echo=FALSE}
library(forecast)

# ACF Plot
ggAcf(df$W398RC1A027NBEA) +
  labs(title = "ACF of Household Income Time Series") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

# PACF Plot
ggPacf(df$W398RC1A027NBEA) +
  labs(title = "PACF of Household Income Time Series") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

```

From here, we can see that the dataset has a correlation, which also means that it is not stationary. In order to prove this conclusion, we utilize the adf test to ensure the results are the same.


### Validation with ADF Test
```{r, echo=FALSE}
library(tseries)
adf.test(df$W398RC1A027NBEA)


```

Based on the result, we can see that the p-value is greater than the thershold value, this means that we fail to reject the Null hypothesis. The time series dataset is not stationary.

Then, to explore more, we utilize detrended and log-transformation to see about the patterns


### Detrended and Log-transformed
```{r, echo=FALSE}
detrended_data <- data.frame(Date = df$DATE[-1], Detrended = diff(df$W398RC1A027NBEA))

# Enhanced Detrended Plot
ggplot(data = detrended_data, aes(x = Date, y = Detrended)) +
  geom_line(color = "blue") +
  labs(title = "Detrended Household Income Time Series",
       x = "Date",
       y = "Detrended Value") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

```


```{r,warning=FALSE, echo=FALSE}
log_transformed_data <- data.frame(Date = df$DATE, LogTransformed = log(df$W398RC1A027NBEA))

# Enhanced Log-transformed Plot with Custom Background
ggplot(data = log_transformed_data, aes(x = Date, y = LogTransformed)) +
  geom_line(color = "blue") +
  labs(title = "Log-transformed Time Series of Household Income",
       x = "Date",
       y = "Log-transformed Value") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

```

From here, we can see that after detrending and dfferencing, the household income remains fluctuations espectially during 2020 period, which could be the cause of the Covid-19. In addition, Log-transformed time series, in order to remove the hetroscadastisity, we can see that the trend remains. I think the reason is possibly because log-transformation can assuage issues related to non-constant variance, which means that the inherent trend within the data may persist.

# Test it stationary again
```{r, echo=FALSE}
library(tseries)
adf.test(log_transformed_data$LogTransformed)


```
```{r}
fit = lm(df$W398RC1A027NBEA~time(df$W398RC1A027NBEA), na.action=NULL) 
# regress chicken on time
#time creates the vector of times at which a time series was sampled.

summary(fit)
```

### Make it Stationary by using differencing

```{r}
t = ts(diff(diff(df$W398RC1A027NBEA)))
# Create the time series plot with the desired background and borders
ts_plot <- autoplot(t) +
  labs(title = "Time Series Plot With two Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Extract the ACF and PACF plots without the theme
acf_plot <- ggAcf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "ACF for two Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )
pacf_plot <- ggPacf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "PACF for two Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Combine the plots
grid.arrange(ts_plot, acf_plot, pacf_plot, ncol = 1)

```

```{r}
library(gridExtra)
library(forecast)
library(ggplot2)
t = ts(diff(diff(diff(df$W398RC1A027NBEA))))
# Create the time series plot with the desired background and borders
ts_plot <- autoplot(t) +
  labs(title = "Time Series Plot With three Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Extract the ACF and PACF plots without the theme
acf_plot <- ggAcf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "ACF for three Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )
pacf_plot <- ggPacf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "PACF for three Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Combine the plots
grid.arrange(ts_plot, acf_plot, pacf_plot, ncol = 1)

```



```{r, warning=FALSE}
adf.test(diff(df$W398RC1A027NBEA, differences = 3))


```

```{r, warning=FALSE}
# Load required libraries
library(ggplot2)
library(forecast)

# First, perform the augmented Dickey-Fuller test
adf_result <- adf.test(diff(df$W398RC1A027NBEA, differences = 3))

# Print the ADF test result
print(adf_result)

# Set the theme
theme_demo <- theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

# Plot the ACF and PACF of the differenced series with the demo theme
par(mfrow=c(2,1)) # Set up a 2x1 grid for plotting
diff_series = diff(df$W398RC1A027NBEA, differences = 3)
# ACF Plot
acf_plot <- acf(diff_series, plot = FALSE)
print(acf_plot, main="ACF of Differenced Household Income Time Series")
ggAcf(diff_series) + theme_demo

# PACF Plot
pacf_plot <- pacf(diff_series, plot = FALSE)
print(pacf_plot, main="PACF of Differenced Household Income Time Series")
ggPacf(diff_series) + theme_demo

```

From the PACF plot, it seems that after the 1st lag, the correlation values drop into insignificance. Thus, we might consider p = 1 for the AR component. From the ACF plot you provided, the correlation seems to cut off after the 1st lag as well. Hence, you might consider q = 1 for the MA component. And we use d = 3 since we utilized third orders of differencing.

# Fit model
```{r}
library(forecast)
diff = diff(df$W398RC1A027NBEA, differences = 3)
model <- Arima(diff, order=c(1,3,2))
summary(model)

```

# Diagnotis

```{r,warning=FALSE}
## empty list to store model fits
set.seed((150))
arma14 =arima.sim(list(order=c(1,3,1), ar=-.4, ma=c(-.3)), n=10000)
arma14 %>% ggtsdisplay()
ARMA_res <- list()

## set counter
cc <-1

## loop over AR
for(p in 0:3){
  ## loop over MA
  for(q in 0:4){
    
  ARMA_res[[cc]]<-arima(x=arma14,order = c(p,3,q))
  cc<- cc+1
  }
}

## get AIC values for model evaluation

ARMA_AIC<-sapply(ARMA_res,function(x) x$aic)

ARMA_res[[which(ARMA_AIC == min(ARMA_AIC))]]
```

```{r,warning=FALSE}
d=3
i=1
temp= data.frame()
ls=matrix(rep(NA,6*23),nrow=23) # roughly nrow = 3x4x2


for (p in 3:5)# p=1,2,3 : 3
{
  for(q in 3:5)# q=1,2,3,4 :4
  {
    for(d in 1:3)# d=1,2 :2
    {
      
      if(p-1+d+q-1<=8) #usual threshold
      {
        
        model<- Arima(df$W398RC1A027NBEA,order=c(p-1,d,q-1),include.drift=TRUE) 
        ls[i,]= c(p-1,d,q-1,model$aic,model$bic,model$aicc)
        i=i+1
        #print(i)
        
      }
      
    }
  }
}

temp= as.data.frame(ls)
names(temp)= c("p","d","q","AIC","BIC","AICc")

#temp
knitr::kable(temp)
```
```{r}
temp[which.min(temp$AIC),]
```
```{r}
temp[which.min(temp$BIC),]
```
```{r}
temp[which.min(temp$AICc),]
```



```{r,warning=FALSE}
######### compare 3 models
set.seed(345)
model_output <- capture.output(sarima(df$W398RC1A027NBEA, 2,3,2))
```

```{r}
model_output2 <- capture.output(sarima(df$W398RC1A027NBEA, 1,3,2))
```

```{r,warning=FALSE}
#################### fitted vs. actual plot  ##############

fit1=Arima(df$W398RC1A027NBEA,order=c(1,3,2),include.drift = TRUE)
fit2=Arima(df$W398RC1A027NBEA,order=c(2,3,2),include.drift = TRUE) #warning for the drift, no drift 


# Set background color
par(bg = "#E0E0E0")

# Plotting the data and model fits with custom background and border colors
plot(df$W398RC1A027NBEA, col="blue", main="Time Series with SARIMA Fits", 
     ylab="Value", xlab="Time", 
     panel.first = rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], 
                        col = "#E0E0E0", border = "#E0E0E0"))

lines(fitted(fit1), col="green")
lines(fitted(fit2), col="red") 

# Adding a legend
legend("topleft", 
       legend = c("The Time Series Of the Data", "fit1", "fit2"), 
       col = c("blue", "green", "red"), 
       lty=1, # Type of the line (1 is for "solid")
       cex=0.8, # Font size of the legend text
       box.col="#E0E0E0", bg="#E0E0E0") 

```

```{r}
forecast(fit2,10) # Forecasting for the next 10 minutes
```

```{r}


# Autoplot with custom colors
plot_fit2 <- autoplot(forecast(fit2)) + 
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    legend.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    legend.key = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Display the plot
print(plot_fit2)
```
# BENCHMARK

```{r}
library(forecast)
diff = diff(df$W398RC1A027NBEA, differences = 3)
model <- Arima(diff, order=c(1,3,2))

```

```{r}
f1<-meanf(df$W398RC1A027NBEA, h=11) #mean

checkresiduals(f1)
```

```{r}
df$W398RC1A027NBEA %>%
  Arima(order=c(0,0,1), seasonal=c(0,1,1)) %>% 
  residuals() %>% 
  ggtsdisplay()
```

```{r}
fit=Arima(df$W398RC1A027NBEA, order = c(0,1,1),seasonal = list(order=c(0,0,1), period=4) ) 
summary(fit)
```

```{r}
f1 <- meanf(df$W398RC1A027NBEA, h=10) 

#checkresiduals(f1)

f2 <- naive(df$W398RC1A027NBEA, h=10) 

#checkresiduals(f2)

f3 <- rwf(df$W398RC1A027NBEA,drift=TRUE, h=10) 
```

```{r}
pred=forecast(fit2,20)
accuracy(pred)
pred[['mean']]
```

```{r}
accuracy(f1)
```

```{r}
accuracy(f2)
```

```{r}
accuracy(f3)
```



```{r}
gdp.ts=ts(df$W398RC1A027NBEA)
```

```{r}
autoplot(gdp.ts) +
  autolayer(meanf(gdp.ts, h=11),
            series="Mean", PI=FALSE) +
  autolayer(naive(gdp.ts, h=11),
            series="Naïve", PI=FALSE) +
  autolayer(snaive(gdp.ts, h=11),
            series="fit", PI=FALSE) +
  autolayer(forecast(fit2, h=11),
            series="Seasonal naïve", PI=FALSE) +
  ggtitle("Forecasts for Household Income") +
  xlab("Year") + ylab("Household Income") +
  guides(colour=guide_legend(title="Forecast")) +
  theme(
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    panel.background = element_rect(fill = "#E0E0E0"),
    panel.border = element_rect(fill = NA, color = "#E0E0E0")
  )
```

```{r}
autoplot(gdp.ts) +
  autolayer(meanf(gdp.ts, h=20),
            series="Mean.tr", PI=FALSE) +
  autolayer(naive(gdp.ts, h=20),
            series="Naïve.tr", PI=FALSE) +
  autolayer(rwf(gdp.ts, drift=TRUE, h=40),
            series="Drift.tr", PI=FALSE) +
  autolayer(forecast(fit2,20), 
            series="fit",PI=FALSE) +
  ggtitle("Household Income in US)") +
  xlab("Time") + ylab("Household Income") +
  guides(colour=guide_legend(title="Forecast")) +
  theme(
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    panel.background = element_rect(fill = "#E0E0E0"),
    panel.border = element_rect(fill = NA, color = "#E0E0E0")
  )
```

