---
title: "datavis"
author: "Zonghong Yu"
date: "2023-09-18"
output: html_document
---

```{r}
library(reticulate)
library(ggplot2)
library(forecast)
library(astsa) 
library(gridExtra)

library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(lubridate)
library(plotly)
library(TTR) # For the SMA function

```


```{r, echo=FALSE, results='hide',warning=FALSE,message=FALSE}
df <- read.csv("../../Dataset/project/household_saving.csv")
df$DATE <- as.Date(df$DATE)
head(df)
```

```{r,warning=FALSE,echo=FALSE}
library(ggplot2)


ggplot(data = df, aes(x = DATE, y = W398RC1A027NBEA)) +
  geom_line(color = "DarkViolet") +
  labs(title = "Time Series Plot of Household Saving",
       x = "Date", 
       y = "The Household Saving") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))


```






### ACF & PACF
```{r, echo=FALSE}
library(forecast)

# ACF Plot
ggAcf(df$W398RC1A027NBEA) +
  labs(title = "ACF of Household Saving Time Series") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

# PACF Plot
ggPacf(df$W398RC1A027NBEA) +
  labs(title = "PACF of Household Saving Time Series") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

```

From here, we can see that the dataset has a correlation, which also means that it is not stationary. In order to prove this conclusion, we utilize the adf test to ensure the results are the same.


### Validation with ADF Test
```{r, echo=FALSE}
library(tseries)
adf.test(df$W398RC1A027NBEA)


```

Based on the result, we can see that the p-value is greater than the thershold value, this means that we fail to reject the Null hypothesis. The time series dataset is not stationary.

Then, to explore more, we utilize detrended and log-transformation to see about the patterns


### Detrended and Log-transformed
```{r, echo=FALSE}
detrended_data <- data.frame(Date = df$DATE[-1], Detrended = diff(df$W398RC1A027NBEA))

# Enhanced Detrended Plot
ggplot(data = detrended_data, aes(x = Date, y = Detrended)) +
  geom_line(color = "blue") +
  labs(title = "Detrended Household Saving Time Series",
       x = "Date",
       y = "Detrended Value") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

```


```{r,warning=FALSE, echo=FALSE}
log_transformed_data <- data.frame(Date = df$DATE, LogTransformed = log(df$W398RC1A027NBEA))

# Enhanced Log-transformed Plot with Custom Background
ggplot(data = log_transformed_data, aes(x = Date, y = LogTransformed)) +
  geom_line(color = "blue") +
  labs(title = "Log-transformed Time Series of Household Saving",
       x = "Date",
       y = "Log-transformed Value") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

```

From here, we can see that after detrending and dfferencing, the household Saving remains fluctuations espectially during 2020 period, which could be the cause of the Covid-19. In addition, Log-transformed time series, in order to remove the hetroscadastisity, we can see that the trend remains. I think the reason is possibly because log-transformation can assuage issues related to non-constant variance, which means that the inherent trend within the data may persist.

# Test it stationary again
```{r, echo=FALSE}
library(tseries)
adf.test(log_transformed_data$LogTransformed)


```
```{r}
fit = lm(df$W398RC1A027NBEA~time(df$W398RC1A027NBEA), na.action=NULL) 
# regress chicken on time
#time creates the vector of times at which a time series was sampled.

summary(fit)
```

### Make it Stationary by using differencing

```{r}

t = ts(diff(diff(df$W398RC1A027NBEA)))
# Create the time series plot with the desired background and borders
ts_plot <- autoplot(t) +
  labs(title = "Time Series Plot With two Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Extract the ACF and PACF plots without the theme
acf_plot <- ggAcf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "ACF for two Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )
pacf_plot <- ggPacf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "PACF for two Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Combine the plots
grid.arrange(ts_plot, acf_plot, pacf_plot, ncol = 1)

```

```{r}
library(gridExtra)
library(forecast)
library(ggplot2)
t = ts(diff(diff(diff(df$W398RC1A027NBEA))))
# Create the time series plot with the desired background and borders
ts_plot <- autoplot(t) +
  labs(title = "Time Series Plot With three Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Extract the ACF and PACF plots without the theme
acf_plot <- ggAcf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "ACF for three Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )
pacf_plot <- ggPacf(diff(diff(diff(df$W398RC1A027NBEA)))) +
  labs(title = "PACF for three Orders of Differencing") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Combine the plots
grid.arrange(ts_plot, acf_plot, pacf_plot, ncol = 1)

```



```{r, warning=FALSE}
adf.test(diff(df$W398RC1A027NBEA, differences = 3))


```

```{r, warning=FALSE}
# Load required libraries
library(ggplot2)
library(forecast)

# First, perform the augmented Dickey-Fuller test
adf_result <- adf.test(diff(df$W398RC1A027NBEA, differences = 3))

# Print the ADF test result
print(adf_result)

# Set the theme
theme_demo <- theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

# Plot the ACF and PACF of the differenced series with the demo theme
par(mfrow=c(2,1)) # Set up a 2x1 grid for plotting
diff_series = diff(df$W398RC1A027NBEA, differences = 3)
# ACF Plot
acf_plot <- acf(diff_series, plot = FALSE)
print(acf_plot, main="ACF of Differenced Household Saving Time Series")
ggAcf(diff_series) + theme_demo

# PACF Plot
pacf_plot <- pacf(diff_series, plot = FALSE)
print(pacf_plot, main="PACF of Differenced Household Saving Time Series")
ggPacf(diff_series) + theme_demo

```

From the PACF plot, it seems that after the 1st lag, the correlation values drop into insignificance. Thus, we might consider p = 1 for the AR component. From the ACF plot you provided, the correlation seems to cut off after the 1st lag as well. Hence, you might consider q = 1 for the MA component. And we use d = 3 since we utilized third orders of differencing.

# Fit model
```{r}
library(forecast)
diff = diff(df$W398RC1A027NBEA, differences = 3)
model <- Arima(diff, order=c(1,3,2))
summary(model)

```

# Diagnotis

```{r,warning=FALSE}
## empty list to store model fits
set.seed((150))
arma14 =arima.sim(list(order=c(1,3,1), ar=-.4, ma=c(-.3)), n=10000)
arma14 %>% ggtsdisplay()
ARMA_res <- list()

## set counter
cc <-1

## loop over AR
for(p in 0:3){
  ## loop over MA
  for(q in 0:4){
    
  ARMA_res[[cc]]<-arima(x=arma14,order = c(p,3,q))
  cc<- cc+1
  }
}

## get AIC values for model evaluation

ARMA_AIC<-sapply(ARMA_res,function(x) x$aic)

ARMA_res[[which(ARMA_AIC == min(ARMA_AIC))]]
```

```{r,warning=FALSE}
d=3
i=1
temp= data.frame()
ls=matrix(rep(NA,6*23),nrow=23) # roughly nrow = 3x4x2


for (p in 3:5)# p=1,2,3 : 3
{
  for(q in 3:5)# q=1,2,3,4 :4
  {
    for(d in 1:3)# d=1,2 :2
    {
      
      if(p-1+d+q-1<=8) #usual threshold
      {
        
        model<- Arima(df$W398RC1A027NBEA,order=c(p-1,d,q-1),include.drift=TRUE) 
        ls[i,]= c(p-1,d,q-1,model$aic,model$bic,model$aicc)
        i=i+1
        #print(i)
        
      }
      
    }
  }
}

temp= as.data.frame(ls)
names(temp)= c("p","d","q","AIC","BIC","AICc")

#temp
knitr::kable(temp)
```
```{r}
temp[which.min(temp$AIC),]
```
```{r}
temp[which.min(temp$BIC),]
```
```{r}
temp[which.min(temp$AICc),]
```



```{r,warning=FALSE}
######### compare 3 models
set.seed(345)
model_output <- capture.output(sarima(df$W398RC1A027NBEA, 2,3,2))
```

```{r}
model_output2 <- capture.output(sarima(df$W398RC1A027NBEA, 1,3,2))
```

```{r,warning=FALSE}
#################### fitted vs. actual plot  ##############

fit1=Arima(df$W398RC1A027NBEA,order=c(1,3,2),include.drift = TRUE)
fit2=Arima(df$W398RC1A027NBEA,order=c(2,3,2),include.drift = TRUE) #warning for the drift, no drift 


# Set background color
par(bg = "#E0E0E0")

# Plotting the data and model fits with custom background and border colors
plot(df$W398RC1A027NBEA, col="blue", main="Time Series with SARIMA Fits", 
     ylab="Value", xlab="Time", 
     panel.first = rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], 
                        col = "#E0E0E0", border = "#E0E0E0"))

lines(fitted(fit1), col="green")
lines(fitted(fit2), col="red") 

# Adding a legend
legend("topleft", 
       legend = c("The Time Series Of the Data", "fit1", "fit2"), 
       col = c("blue", "green", "red"), 
       lty=1, # Type of the line (1 is for "solid")
       cex=0.8, # Font size of the legend text
       box.col="#E0E0E0", bg="#E0E0E0") 

```

```{r}
forecast(fit2,10) # Forecasting for the next 10 minutes
```

```{r}


# Autoplot with custom colors
plot_fit2 <- autoplot(forecast(fit2)) + 
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    legend.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    legend.key = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Display the plot
print(plot_fit2)
```
# BENCHMARK

```{r}
library(forecast)
diff = diff(df$W398RC1A027NBEA, differences = 3)
model <- Arima(diff, order=c(1,3,2))

```

```{r}
f1<-meanf(df$W398RC1A027NBEA, h=11) #mean

checkresiduals(f1)
```

```{r}
df$W398RC1A027NBEA %>%
  Arima(order=c(0,0,1), seasonal=c(0,1,1)) %>% 
  residuals() %>% 
  ggtsdisplay()
```

```{r}
fit=Arima(df$W398RC1A027NBEA, order = c(0,1,1),seasonal = list(order=c(0,0,1), period=4) ) 
summary(fit)
```

```{r}
f1 <- meanf(df$W398RC1A027NBEA, h=10) 

#checkresiduals(f1)

f2 <- naive(df$W398RC1A027NBEA, h=10) 

#checkresiduals(f2)

f3 <- rwf(df$W398RC1A027NBEA,drift=TRUE, h=10) 
```

```{r}
pred=forecast(fit2,20)
accuracy(pred)
pred[['mean']]
```

```{r}
accuracy(f1)
```

```{r}
accuracy(f2)
```

```{r}
accuracy(f3)
```



```{r}
gdp.ts=ts(df$W398RC1A027NBEA)
```

```{r}
autoplot(gdp.ts) +
  autolayer(meanf(gdp.ts, h=11),
            series="Mean", PI=FALSE) +
  autolayer(naive(gdp.ts, h=11),
            series="Naïve", PI=FALSE) +
  autolayer(snaive(gdp.ts, h=11),
            series="fit", PI=FALSE) +
  autolayer(forecast(fit2, h=11),
            series="Seasonal naïve", PI=FALSE) +
  ggtitle("Forecasts for Household Saving") +
  xlab("Year") + ylab("Household Saving") +
  guides(colour=guide_legend(title="Forecast")) +
  theme(
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    panel.background = element_rect(fill = "#E0E0E0"),
    panel.border = element_rect(fill = NA, color = "#E0E0E0")
  )
```

```{r}
autoplot(gdp.ts) +
  autolayer(meanf(gdp.ts, h=20),
            series="Mean.tr", PI=FALSE) +
  autolayer(naive(gdp.ts, h=20),
            series="Naïve.tr", PI=FALSE) +
  autolayer(rwf(gdp.ts, drift=TRUE, h=40),
            series="Drift.tr", PI=FALSE) +
  autolayer(forecast(fit2,20), 
            series="fit",PI=FALSE) +
  ggtitle("Household Saving in US)") +
  xlab("Time") + ylab("Household Saving") +
  guides(colour=guide_legend(title="Forecast")) +
  theme(
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    panel.background = element_rect(fill = "#E0E0E0"),
    panel.border = element_rect(fill = NA, color = "#E0E0E0")
  )
```



# SARIMA ANALYSIS
# Before Covid
```{r}
df$W398RC1A027NBEA[1:28] 
```

```{r}
USgas <- ts(df$W398RC1A027NBEA[1:28], frequency = 2)
USgas
```

```{r}
autoplot(USgas)+ggtitle("Household Saving") 
```

# Check

```{r}
dec=decompose(USgas,type = c("additive", "multiplicative"))
plot(dec)
```

```{r}
gglagplot(USgas, do.lines=FALSE, set.lags = c(1, 2, 3, 4))
```

Most spikes are within range, it is weakly stationary. 
WE can continue analysis



```{r}
# Create the time series plot with the desired background and borders
ts_plot <- autoplot(USgas) +
  labs(title = "Time Series Plot ") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Extract the ACF and PACF plots without the theme
acf_plot <- ggAcf(USgas) +
  labs(title = "ACF for data") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )
pacf_plot <- ggPacf(USgas) +
  labs(title = "PACF for data") +
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Combine the plots
grid.arrange(ts_plot, acf_plot, pacf_plot, ncol = 1)

```


From previous analysis with ARIMA, we used third order differencing due to the adf test. however, we should know that the adf test result is not as reliable as pacf plots. Here, we can see that the data is already weakly stationary. Therefore, we can choose the corresponind parameter sets:


The ACF (Autocorrelation Function) and PACF (Partial Autocorrelation Function) plots are commonly used for identifying the order of the ARIMA or SARIMA models.

From the plots:

- **ACF** is showing no significant spikes after lag 1,2,3,4 which suggests there isn't too many significant autocorrelations in the series.
  
- **PACF** has a significant spike at lag 1 and then quickly drops, suggesting an AR term of order 1.

For a non-seasonal ARIMA model (p,d,q):

- **p (AR order)**: Based on PACF, it seems like p=1 might be a good choice.
  
- **d (differencing order)**: We can start with d=0 or d=1.
  
- **q (MA order)**: Based on the ACF, it seems like q=0 or q=1 might be a good choice.

For a seasonal SARIMA model (P,D,Q):

- **P (seasonal AR order)**: We'd typically look for a spike at the seasonal lag in the PACF. Since there doesn't appear to be a clear seasonal pattern, we might start with P=0,1.
  
- **D (seasonal differencing order)**: Similarly, without clear seasonality, you might start with D=0. We'd look for seasonality in the original data to determine this.
  
- **Q (seasonal MA order)**: We'd look for a spike at the seasonal lag in the ACF. Again, without a clear seasonal pattern, Q=0,1 might be a starting point.


```{r}
######################## Check for different combinations ########
SARIMA.c=function(p1,p2,q1,q2,P1,P2,Q1,Q2,data){
  temp=c()
  d=0
  D=0
  s=0
  
  i=1
  temp= data.frame()
  ls=matrix(rep(NA,9*25),nrow=25)
  
  for (p in p1:p2)
  {
    for(q in q1:q2)
    {
      for(P in P1:P2)
      {
        for(Q in Q1:Q2)
        {
          if(p+d+q+P+D+Q<=9)
          {
            model<- Arima(data,order=c(p-1,d,q-1),seasonal=c(P-1,D,Q-1))
            ls[i,]= c(p-1,d,q-1,P-1,D,Q-1,model$aic,model$bic,model$aicc)
            i=i+1
          }
        }
      }
    }
  }
  
  temp= as.data.frame(ls)
  names(temp)= c("p","d","q","P","D","Q","AIC","BIC","AICc")
  temp
}


  

```

```{r}
# Based on the analysis:

output=SARIMA.c(p1=1,p2=2,q1=1,q2=3,P1=1,P2=2,Q1=1,Q2=2,data=USgas)
knitr::kable(output)

```


```{r}
output[which.min(output$AIC),] 
```

```{r}
output[which.min(output$BIC),]
```

```{r}
output[which.min(output$AICc),]
```


```{r}

set.seed(236)
model_output <- capture.output(sarima(USgas, 1,0,0,0,0,0,0))
```


```{r}
cat(model_output[88:112], model_output[length(model_output)], sep = "\n") 
```

The Standard Residual Plot appears good, displaying okay stationarity with a nearly constant mean and variation.

The Autocorrelation Function (ACF) plot shows almost no correlation indicating that the model has harnessed everything that left is white noise. This indicates a good model fit.

The Quantile-Quantile (Q-Q) Plot still demonstrates near-normality.

The Ljung-Box test results reveal values below the 0.05 (5% significance) threshold, indicating there’s some significant correlation left.

$ttable: all coefficients are significant.



# Fit model

```{r}
fit <- Arima(USgas, order=c(1,0,0), seasonal=c(1,0,0))
summary(fit)
```


```{r}
# Autoplot with custom colors
plot_fit_ <- autoplot(forecast(fit,12)) + 
  theme(
    panel.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    plot.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    legend.background = element_rect(fill = "#E0E0E0", color = "#E0E0E0"),
    legend.key = element_rect(fill = "#E0E0E0", color = "#E0E0E0")
  )

# Display the plot
print(plot_fit_)

```

Since that we do not have certain seasonality, it is reasonable that the model did not capture any seasonal pattern. The model captured the downwarding trend based on the end of the dataset.

```{r}
sarima.for(USgas, 36, 1,0,0,1,1,0,12)
```


```{r}

fit <- Arima(USgas, order=c(1,1,1), seasonal=c(0,1,1))

autoplot(USgas) +
  autolayer(meanf(USgas, h=36),
            series="Mean", PI=FALSE) +
  autolayer(naive(USgas, h=36),
            series="Naïve", PI=FALSE) +
  autolayer(snaive(USgas, h=36),
            series="SNaïve", PI=FALSE)+
  autolayer(rwf(USgas, h=36, drift=TRUE),
            series="Drift", PI=FALSE)+
  autolayer(forecast(fit,36), 
            series="fit",PI=FALSE) +
  guides(colour=guide_legend(title="Forecast"))
```

```{r}
f2 <- snaive(USgas, h=36) 

accuracy(f2)
```

```{r}
summary(fit)
```

Although I concluded that the SARIMA did perform too well since we did not have seasonal pattern, it is clear that it outperforms the other benchmark methods.
