---
title: "datavis"
author: "Zonghong Yu"
date: "2023-09-18"
output: html_document
---

```{r}
library(reticulate)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(lubridate)
library(plotly)
library(TTR) # For the SMA function

```


```{r, echo=FALSE, results='hide',warning=FALSE,message=FALSE}
df <- read.csv("../../Dataset/project/household_saving.csv")
df$DATE <- as.Date(df$DATE)
head(df)
```

```{r,warning=FALSE,echo=FALSE}
library(ggplot2)


ggplot(data = df, aes(x = DATE, y = W398RC1A027NBEA)) +
  geom_line(color = "DarkViolet") +
  labs(title = "Time Series Plot of Household Income",
       x = "Date", 
       y = "The Household Income") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))


```
```{r}
# Calculate Moving Averages
df$MA2 <- SMA(df$W398RC1A027NBEA, n=2)
df$MA4 <- SMA(df$W398RC1A027NBEA, n=4)
df$MA8 <- SMA(df$W398RC1A027NBEA, n=8)

```


```{r}

fig <- df %>%
  plot_ly(x = ~DATE, type = "scatter", mode = "lines") %>%
  add_lines(y = ~W398RC1A027NBEA, name = 'Original', line = list(color = 'blue')) %>%
  add_lines(y = ~MA2, name = 'MA(2)', line = list(color = 'red', dash = 'dash')) %>%
  add_lines(y = ~MA4, name = 'MA(4)', line = list(color = 'green', dash = 'dash')) %>%
  add_lines(y = ~MA8, name = 'MA(8)', line = list(color = 'purple', dash = 'dash')) %>%
  layout(title = "Moving Average Smoothing",
         xaxis = list(title = "Date"),
         yaxis = list(title = "W398RC1A027NBEA"),
         plot_bgcolor = '#E0E0E0',  # background color of the plotting area
         paper_bgcolor = '#E0E0E0', # background color of the entire plot (around the plotting area)
         legend = list(title = list(text = "W398RC1A027NBEA Data"),
                       itemsizing = "constant")
         )

# Display the plot
fig

```


```{r}
# Assuming df has been loaded as per your snippet

# Load necessary library
library(ggplot2)

# MA function to calculate moving averages
ma_function <- function(data, window){
  stats::filter(data, rep(1/window, window), sides=2)
}

# Window sizes for moving average
windows <- c(2, 4, 8)  # Example windows, you can adjust them as per your needs

# Create a plot
ggplot(df, aes(x=DATE)) +
  geom_line(aes(y=W398RC1A027NBEA), colour="blue", linetype="dotted") +
  geom_line(aes(y=ma_function(W398RC1A027NBEA, windows[1])), colour="red") +
  geom_line(aes(y=ma_function(W398RC1A027NBEA, windows[2])), colour="green") +
  geom_line(aes(y=ma_function(W398RC1A027NBEA, windows[3])), colour="purple") +
  labs(title="Moving Average Smoothing",
       subtitle="Using window sizes: 2 (red), 4 (green), 8 (purple)",
       y="W398RC1A027NBEA",
       x="Date") +
  theme_minimal()


```
```{r}

library(plotly)

```

```{r}
ma_function <- function(data, window){
  stats::filter(data, rep(1/window, window), sides=2)
}

```

```{r}
windows <- c(5, 10, 20)

ma_2 <- ma_function(df$W398RC1A027NBEA, windows[1])
ma_4 <- ma_function(df$W398RC1A027NBEA, windows[2])
ma_8 <- ma_function(df$W398RC1A027NBEA, windows[3])

```

```{r}
p <- plot_ly(df, x = ~DATE, type='scatter', mode='lines') %>%
  add_lines(y = ~W398RC1A027NBEA, name='Original', line=list(color='blue')) %>%
  add_lines(y = ~ma_2, name='MA(2)', line=list(color='red')) %>%
  add_lines(y = ~ma_4, name='MA(4)', line=list(color='green')) %>%
  add_lines(y = ~ma_8, name='MA(8)', line=list(color='purple')) %>%
  layout(title="Moving Average Smoothing",
         xaxis=list(title="Date"),
         yaxis=list(title="W398RC1A027NBEA"))

# Show plot
p

```


```{r}
df <- read.csv("../../Dataset/project/household_saving.csv")
df$DATE <- as.Date(df$DATE)
head(df)
```

```{r, echo=FALSE, results='hide',warning=FALSE,message=FALSE}
df <- read.csv("../../Dataset/project/A191RI1Q225SBEA.csv")
df$DATE <- as.Date(df$DATE)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}

library(plotly)



# Create a plotly plot with background color set to #E0E0E0 and line color set to grey
p <- plot_ly(data = df, x = ~DATE, y = ~A191RI1Q225SBEA, type = 'scatter', mode = 'lines', 
             line = list(color = 'grey')) %>%
  layout(
    title = "GDP Over Time In USA",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Percent Change from Preceding Period"),
    plot_bgcolor = '#E0E0E0',   # Setting the background color
    paper_bgcolor = '#E0E0E0'  # Setting the surrounding background color
  )

# Display the plot
p

```   


```{r, echo=FALSE}
library(htmlwidgets)
saveWidget(p, file = "../../image/DataSource/gdp.html")

```





# GDP EDA Section

```{r,warning=FALSE}
library(ggplot2)


ggplot(data = df, aes(x = DATE, y = A191RI1Q225SBEA)) +
  geom_line(color = "blue") +
  labs(title = "Time Series Plot of GDP",
       x = "Date", 
       y = "Percent Change from Preceding Period") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))


```
From the plot, you can visually inspect:

Trend: Any consistent upward or downward direction.
Seasonality: Any repeating patterns or cycles.
Variation: Fluctuations in the data.
Periodic fluctuations: Spikes or drops at consistent intervals.
Determine if the time series looks multiplicative or additive. An additive time series has constant amplitude and frequency, while a multiplicative one has varying amplitude/frequency.

```{r}
# Lag plot
lagged_data <- data.frame(value = df$A191RI1Q225SBEA[-1],
                          lagged_value = df$A191RI1Q225SBEA[-length(df$A191RI1Q225SBEA)])

# Enhanced Lag Plot
ggplot(data = lagged_data, aes(x = lagged_value, y = value)) +
  geom_point(color = "blue", alpha = 0.5) +
  labs(title = "Lag Plot (lag=1)",
       x = "Value at t-1",
       y = "Value at t") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))



```
You'd interpret the lag plot by looking for any structure. If the points cluster along a diagonal line from bottom-left to top-right, there's a positive correlation. Any other pattern might suggest non-linearity or some pattern not captured by mere linear correlation

```{r}
#library(ggfortify)

# Decomposition using ggplot2 styling
decomposed <- decompose(ts(df$A191RI1Q225SBEA, frequency=12), type = "additive")
autoplot(decomposed) + 
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

```
```{r}
library(forecast)

# ACF Plot
ggAcf(df$A191RI1Q225SBEA) +
  labs(title = "ACF of GDP Time Series") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

# PACF Plot
ggPacf(df$A191RI1Q225SBEA) +
  labs(title = "PACF of GDP Time Series") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        plot.background = element_rect(fill = "#E0E0E0"))

```
The ACF plot can help determine if the series is stationary. If the series is non-stationary, the ACF will decay slowly, while for a stationary series, the ACF will drop off quickly

```{r}

# ACF and PACF
acf(df$A191RI1Q225SBEA)
pacf(df$A191RI1Q225SBEA)

```

```{r}
library(tseries)
adf.test(df$A191RI1Q225SBEA)


```

stationary !

```{r}
detrended_data <- data.frame(Date = df$DATE[-1], Detrended = diff(df$A191RI1Q225SBEA))

# Enhanced Detrended Plot
ggplot(data = detrended_data, aes(x = Date, y = Detrended)) +
  geom_line(color = "blue") +
  labs(title = "Detrended Time Series",
       x = "Date",
       y = "Detrended Value") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

```


```{r}
log_transformed <- log(df$A191RI1Q225SBEA)
plot(log_transformed, main = "Log-transformed Time Series")

```

```{r,warning=FALSE}
log_transformed_data <- data.frame(Date = df$DATE, LogTransformed = log(df$A191RI1Q225SBEA))

# Enhanced Log-transformed Plot with Custom Background
ggplot(data = log_transformed_data, aes(x = Date, y = LogTransformed)) +
  geom_line(color = "blue") +
  labs(title = "Log-transformed Time Series of GDP",
       x = "Date",
       y = "Log-transformed Value") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#E0E0E0"),
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.05),
        plot.background = element_rect(fill = "#E0E0E0"))

```



# Moving Average
```{r}


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
