---
title: "datavis"
author: "Zonghong Yu"
date: "2023-09-18"
output: html_document
---

```{r}
library(reticulate)
library(ggplot2)
library(forecast)
library(astsa) 
library(gridExtra)
library(vars)
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(lubridate)
library(plotly)
library(TTR) # For the SMA function

```


```{r, echo=FALSE, results='hide',warning=FALSE,message=FALSE}
df1 <- read.csv("../../Dataset/project/household_saving.csv")
df2 <- read.csv("../../Dataset/project/MEHOINUSA672N.csv")
df3 <- read.csv("../../Dataset/project/MSPUS.csv")
df4 <- read.csv("../../Dataset/project/SIPOVGINIUSA.csv")
df5 <- read.csv("../../Dataset/project/FIXHAI.csv")
df6 <- read.csv("../../Dataset/project/PSAVERT.csv")
df7 <- read.csv("../../Dataset/project/A191RI1Q225SBEA.csv")
df8 <- read.csv("../../Dataset/project/Merged_downsample.csv")
df9 <- read.csv("../../Dataset/project/USAUCSFRCONDOSMSAMID.csv")
df1$DATE <- as.Date(df1$DATE)
df2$DATE <- as.Date(df2$DATE)
df3$DATE <- as.Date(df3$DATE)
df4$DATE <- as.Date(df4$DATE)
df5$DATE <- as.Date(df5$DATE)
df6$DATE <- as.Date(df6$DATE)
df7$DATE <- as.Date(df7$DATE)
df8$date <- as.Date(df8$date)
df9$DATE <- as.Date(df9$DATE)
head(df9)
```
# Transform all to time series
```{r}
saving =ts(df1$W398RC1A027NBEA)
income =ts(df2$MEHOINUSA672N)
sale = ts(df3$MSPUS)
gini =ts(df4$SIPOVGINIUSA)
afford =ts(df5$FIXHAI)
saverate =ts(df6$PSAVERT)
gdp =ts(df7$A191RI1Q225SBEA)


homevalue =ts(df9$USAUCSFRCONDOSMSAMID)

```



```{r}
length(saving)
length(income)
length(sale)
length(gini)
length(afford)
length(saverate)
length(gdp)
length(homevalue)


```

```{r}
df7_sub$DATE
```

Home Value ~ Saving Rate + Gdp Deflator + Sale Price
```{r}
# Define start and end dates
start_date <- as.Date("2000-01-01")
end_date <- as.Date("2020-12-31")

# Subset data frames to include only data from 1992 through 2020
df3_sub <- subset(df3, DATE >= start_date & DATE <= end_date)
df6_sub <- subset(df6, DATE >= start_date & DATE <= end_date)
df7_sub <- subset(df7, DATE >= start_date & DATE <= end_date)
df9_sub <- subset(df9, DATE >= start_date & DATE <= end_date)

# Assuming the data is annual for this example. If it's quarterly, you'd use frequency = 4; if monthly, frequency = 12.
frequency <- 4

# Create ts objects
sale <- ts(df3_sub$MSPUS, start = c(2000, 1), end = c(2020, 1), frequency = frequency)
saving <- ts(df6_sub$PSAVERT, start = c(2000, 1), end = c(2020, 1), frequency = frequency)
gdp <- ts(df7_sub$A191RI1Q225SBEA, start = c(2000, 1), end = c(2020, 1), frequency = frequency)
homevalue <- ts(df9_sub$USAUCSFRCONDOSMSAMID, start = c(2000, 1), end = c(2020, 1), frequency = frequency)
DATE <- ts(df7_sub$DATE, start = c(2000, 1), end = c(2020, 1), frequency = frequency)
```

```{r}
length(sale)
length(saving )
length(gdp)
length(homevalue)
length(DATE)
```

```{r}
DATE
```


## Model 3 (VAR)
```{r}
dd3<-data.frame(homevalue,saving,gdp,sale,DATE)

colnames(dd3)<-c("homevalue","saving","gdp_deflator", "sale",'date')

knitr::kable(head(dd3))
```

```{r,warning=FALSE}
lg.dd3 <- data.frame("date" =dd3$date,"sale"=log(dd3$sale),"saving"=log(dd3$saving),"homevalue"=log(dd3$homevalue),
                                        "gdp_deflator"=log(dd3$gdp))

#### converting to time series component #########
lg.dd.ts3<-ts(lg.dd3,frequency = 4)

##### Facet Plot #######################
autoplot(lg.dd.ts3[,c(2:5)], facets=TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("The GDP Deflator, Sale Price, Saving Rate, Home value in USA")
```

Fitting a VAR model:
Finding out the best p:
```{r}
VARselect(dd3[, c(2:5)], lag.max=14, type="both")
```
From the results, we can see that 13,14 have relatively smaller criterias:

We can fit several models with p=13, 14.=>  VAR(13), VAR(14)
```{r}
summary(vars::VAR(dd3[, c(2:5)], p=13, type='both'))
```
```{r}
summary(vars::VAR(dd3[, c(2:5)], p=14, type='both'))
```


We can see that some models have some variables significant but not all. We will keep the variables and make comparison in the next steps to get the better one.


## CV

```{r}
n=length(dd3$gdp_deflator)
n*0.3
```
```{r}
k=25 #19*4
n-k
```

```{r}
dd3
```








```{r}
dat = ts(dd3[,c(1,2,3,4)])
```


```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(1)][1:(k-1)+i] 
  xtest <- dat[,c(1)][k+i] 
  
  fit <- vars::VAR(dat, p=13, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=14, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$gdp_deflator-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$gdp_deflator-xtest)^2))

}


```




```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)



# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1+ e, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For Home Value")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 , type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```


```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```

```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(2)][1:(k-1)+i] 
  xtest <- dat[,c(2)][k+i] 
  
  fit <- vars::VAR(dat, p=13, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=14, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$gdp_deflator-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$gdp_deflator-xtest)^2))

}


```




```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)



# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1+ e, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For Saving Rate")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 , type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```


```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```


For gdp deflator
```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(3)][1:(k-1)+i] 
  xtest <- dat[,c(3)][k+i] 
  
  fit <- vars::VAR(dat, p=13, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=14, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$gdp_deflator-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$gdp_deflator-xtest)^2))

}


```



```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)



# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1+ e, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For GDP Deflator")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 , type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```

```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```

```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(4)][1:(k-1)+i] 
  xtest <- dat[,c(4)][k+i] 
  
  fit <- vars::VAR(dat, p=13, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=14, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$gdp_deflator-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$gdp_deflator-xtest)^2))

}


```




```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)



# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1+ e, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For Sale Price")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 , type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```


```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```



The second one VAR(14) is better for Home Value and Sale Price, However, overall, not too different
The first one VAR(13) is better for GDP Deflator and Saving Rate
Therefore, VAR(13) is relatively better than VAR(14)





Forecast
```{r}

fit1 <- vars::VAR(dat[,1:4], p=13, type="const")
fcast1 = predict(fit1, n.ahead = 8, ci = 0.95)
fcast1$fcst$gdp_deflator
```

```{r}
fcast1$fcst$sale
```


```{r}
fcast1$fcst$saving
```

```{r}
fcast1$fcst$homevalue
```

```{r}
fcast1$fcst$sale
```

```{r}
forecast(fit1,48) %>%
  autoplot() + xlab("Year")
```