---
title: "datavis"
author: "Zonghong Yu"
date: "2023-09-18"
output: html_document
---

```{r}
library(reticulate)
library(ggplot2)
library(forecast)
library(astsa) 
library(gridExtra)
library(vars)
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(lubridate)
library(plotly)
library(TTR) # For the SMA function

```


```{r, echo=FALSE, results='hide',warning=FALSE,message=FALSE}
df1 <- read.csv("../../Dataset/project/household_saving.csv")
df2 <- read.csv("../../Dataset/project/MEHOINUSA672N.csv")
df3 <- read.csv("../../Dataset/project/MSPUS.csv")
df4 <- read.csv("../../Dataset/project/SIPOVGINIUSA.csv")
df5 <- read.csv("../../Dataset/project/FIXHAI.csv")
df6 <- read.csv("../../Dataset/project/PSAVERT.csv")
df7 <- read.csv("../../Dataset/project/A191RI1Q225SBEA.csv")
df8 <- read.csv("../../Dataset/project/Merged_downsample.csv")

df1$DATE <- as.Date(df1$DATE)
df2$DATE <- as.Date(df2$DATE)
df3$DATE <- as.Date(df3$DATE)
df4$DATE <- as.Date(df4$DATE)
df5$DATE <- as.Date(df5$DATE)
df6$DATE <- as.Date(df6$DATE)
df7$DATE <- as.Date(df7$DATE)
df8$date <- as.Date(df8$date)
head(df8)
```
# Transform all to time series
```{r}
saving =ts(df1$W398RC1A027NBEA)
income =ts(df2$MEHOINUSA672N)
sale = ts(df3$MSPUS)
gini =ts(df4$SIPOVGINIUSA)
afford =ts(df5$FIXHAI)
saverate =ts(df6$PSAVERT)
gdp =ts(df7$A191RI1Q225SBEA)

saleprice =ts(df8$Mean.Sale.Price)
homevalue =ts(df8$Mean.Home.Value)
rentalprice =ts(df8$mean)
```

```{r}
length(saving)
length(income)
length(sale)
length(gini)
length(afford)
length(saverate)
length(gdp)
length(homevalue)
length(saleprice)
length(rentalprice)

```

saving rate, sale price, gdp deflator
```{r}
# Define start and end dates
start_date <- as.Date("1965-01-01")
end_date <- as.Date("2020-12-31")

# Subset data frames to include only data from 1992 through 2020
df3_sub <- subset(df3, DATE >= start_date & DATE <= end_date)
df6_sub <- subset(df6, DATE >= start_date & DATE <= end_date)
df7_sub <- subset(df7, DATE >= start_date & DATE <= end_date)

# Assuming the data is annual for this example. If it's quarterly, you'd use frequency = 4; if monthly, frequency = 12.
frequency <- 5

# Create ts objects
sale_price <- ts(df3_sub$MSPUS, start = c(1965, 1), end = c(2020, 1), frequency = frequency)
saving_rate <- ts(df6_sub$PSAVERT, start = c(1965, 1), end = c(2020, 1), frequency = frequency)
gdp <- ts(df7_sub$A191RI1Q225SBEA, start = c(1965, 1), end = c(2020, 1), frequency = frequency)
DATE <- ts(df7_sub$DATE, start = c(1965, 1), end = c(2020, 1), frequency = frequency)
```
```{r}
length(sale_price)
length(saving_rate )
length(DATE)
```

```{r}
DATE
```


## Model 3 (VAR)
```{r}
dd3<-data.frame(sale_price,saving_rate,gdp,DATE)

colnames(dd3)<-c("sale_price","saving_rate","gdp_deflator",'date')

knitr::kable(head(dd3))
```

```{r,warning=FALSE}
lg.dd3 <- data.frame("date" =dd3$date,"sale_price"=log(dd3$sale_price),"saving_rate"=log(dd3$saving_rate),
                                        "gdp_deflator"=log(dd3$gdp_deflator))

#### converting to time series component #########
lg.dd.ts3<-ts(lg.dd3,frequency = 5)

##### Facet Plot #######################
autoplot(lg.dd.ts3[,c(2:4)], facets=TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("The GDP Deflator, Sale Price, Saving Rate in USA")
```

Fitting a VAR model:
Finding out the best p:
```{r}
VARselect(dd3[, c(2:4)], lag.max=10, type="both")
```
From the results, we can see that 3,4 have relatively smaller criterias:
We can fit several models with p=1, 3, and 4.=> VAR(1), VAR(3), VAR(4)
```{r}
summary(vars::VAR(dd3[, c(2:4)], p=1, type='both'))
```
```{r}
summary(vars::VAR(dd3[, c(2:4)], p=3, type='both'))
```

```{r}
summary(vars::VAR(dd3[, c(2:4)], p=4, type='both'))
```
We can see that some models have some variables significant but not all. We will keep the variables and make comparison in the next steps to get the better one.


## CV

```{r}
n=length(dd3$gdp_deflator)
n*0.3
```
```{r}
k=83 #19*4
n-k
```

```{r}
dd3
```

```{r}
fit <- vars::VAR(dd3, p=5, type='both')
fcast = predict(fit, n.ahead = 24, ci = 0.95)
mean(fcast$fcst$sale_price)
```

```{r}
fcast$fcst
```


```{r}
n=length(dd3$gdp_deflator)
k=85 #19*4

n*0.3
```



```{r}
dat = ts(dd3[,c(1,2,3)])
```


For gdp deflator
```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(3)][1:(k-1)+i] 
  xtest <- dat[,c(3)][k+i] 
  
  fit <- vars::VAR(dat, p=3, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=4, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$gdp_deflator-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$gdp_deflator-xtest)^2))

}


```



```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)

e <- 0.03

# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1+ e, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For GDP Deflator")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 , type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```

```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```

The first one is better but not too different for GDP Deflator

For saving_rate
```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(2)][1:(k-1)+i] 
  xtest <- dat[,c(2)][k+i] 
  
  fit <- vars::VAR(dat, p=3, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=4, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$saving_rate-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$saving_rate-xtest)^2))

}



```



```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)

e <- 0.03

# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For Saving Rate")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 + e, type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```

Let us reorganize the errors
```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```

The first one is also better for saving rate


## Sale price
```{r}

i=1
err1 = c()
err2 = c()

for(i in 1:(n-k))
{
  xtrain <- dat[,c(1)][1:(k-1)+i] 
  xtest <- dat[,c(1)][k+i] 
  
  fit <- vars::VAR(dat, p=3, type='both')
  fcast1 <- forecast(fit, h=4)
  fcast1 = predict(fit, n.ahead = 12, ci = 0.95)
 
  
  fit2 <- vars::VAR(dat, p=4, type='both')
  fcast2 <- forecast(fit2, h=4)
  fcast2 = predict(fit2, n.ahead = 12, ci = 0.95)
  
  #capture error for each iteration
  # This is RMSE
  err1 = c(err1, sqrt((fcast1$fcst$sale_price-xtest)^2))
  err2 = c(err2, sqrt((fcast2$fcst$sale_price-xtest)^2))

}



```

```{r}
# Normalize function
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize e1 and e2
normalized_e1 <- normalize(err1)
normalized_e2 <- normalize(err2)

e <- 0.03

# Set the background color of the plot region to #E0E0E0
par(bg = "#E0E0E0")

# Plot the normalized e1 with a red line
plot(normalized_e1, type="l", col="red", ylim=range(c(normalized_e1, normalized_e2 + e)), xlab="Index", ylab="Normalized Values", main="Normalized RMSE Plot of e1 and e2 For Sale Price")

# Add the normalized e2 with an offset and a blue line
lines(normalized_e2 + e, type="l", col="blue")

# Add a legend to distinguish the lines
legend("topright", legend=c("e1", "e2"), col=c("red", "blue"), lty=1, cex=0.8)


```

```{r}
error1 <- as.numeric(err1)
error1 <- error1[!is.na(error1) & !is.nan(error1)]
error1 <- mean(error1)

error2 <- as.numeric(err2)
error2 <- error2[!is.na(error2) & !is.nan(error2)]
error2 <- mean(error2)


error1
error2

```

Overall, we can see that in my case, all variables have smaller RMSE with VAR(3)


Forecast
```{r}

fit1 <- vars::VAR(dat[,1:3], p=3, type="const")
fcast1 = predict(fit1, n.ahead = 8, ci = 0.95)
fcast1$fcst$gdp_deflator
```

```{r}
fcast1$fcst$sale_price
```


```{r}
fcast1$fcst$saving_rate
```

```{r}
forecast(fit1,48) %>%
  autoplot() + xlab("Year")
```